Anomalous waves triggered by abrupt depth changes: laboratory experiments and truncated KdV statistical mechanics

### 2019-11-13 paper2_JNLS/paper.tex
NM

%----------------------------------------------------------------------------%
%% LATEX DEFINITIONS
%----------------------------------------------------------------------------%
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\nick}[1]{{\color{red} #1}}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\dx}{\, dx}
\newcommand{\CC}{{\mathbb{C}}}
% Basic physical parameters and scales
\newcommand{\freqp}{f_p}
\newcommand{\etastd}{\eta_{\text std}}
% Parameters that change crossing the ADC
\newcommand{\depth}{d}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\dupdn}{\depth_{\pm}}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamdn}{\lam_{+}}
\newcommand{\lamupdn}{\lam_{\pm}}
\newcommand{\lamfac}{N}
\newcommand{\drat}{\mathcal{D}}
\newcommand{\dratdn}{\drat_+}
\newcommand{\dratupdn}{\drat_{\pm}}
% Statistical quantities
\newcommand{\En}{\mathcal{E}}
\newcommand{\Mo}{\mathcal{M}}
\newcommand{\skw}{\text{skew}}
\newcommand{\skwdn}{\skw_+}
\newcommand{\var}{\text{var}}
\newcommand{\varup}{\var_-}
\newcommand{\vardn}{\var_+}
\newcommand{\varupdn}{\var_{\pm}}
\newcommand{\kurt}{\text{kurt}}
\newcommand{\std}{\text{std}}
% Dimensionless parameters
\newcommand{\ampscale}{\mathcal{A}}
\newcommand{\lengthscale}{\mathcal{L}}
\newcommand{\timescale}{\mathcal{T}}
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}
% Hamiltonian stuff
\newcommand{\uhat}{\hat{u}}
\newcommand{\sympJ}{\mathcal{J}}
\newcommand{\vard}[2]{\frac{\delta #1}{\delta #2}}
\newcommand{\Ham}{\mathcal{H}}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Hupdn}{\Ham^{\pm}}
\newcommand{\Hthree}{\Ham_{3}}
\newcommand{\Htwo}{\Ham_{2}}
\newcommand{\Fcnl}{\mathcal{F}}
% Truncated stuff
\newcommand{\Proj}{\mathcal{P}_{\Lambda}}
\newcommand{\uL}{u_{\Lambda}}
\newcommand{\HLupdn}{\Ham_{\Lambda}^{\pm}}
\newcommand{\SympL}{\sympJ_{\Lambda}}
\newcommand{\uup}{u_{-}}
\newcommand{\udn}{u_{+}}
\newcommand{\uupdn}{u_{\pm}}
% Gibbs and theta
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\Gupdn}{\Gibbs^{\pm}}
\newcommand{\thup}{\theta^{-}}
\newcommand{\thdn}{\theta^{+}}
\newcommand{\thupdn}{\theta^{\pm}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}
\newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
\newcommand{\Gz}{\Gibbs_0}
\newcommand{\meanz}[1]{\mean{#1}_0}
% Stuff
\newcommand{\tavg}[1]{\overline{#1}}
\newcommand{\transf}{\mathcal{F}}
\newcommand{\Nsamp}{N_s}
\newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
\newcommand{\traveldist}{\ell}
\newcommand{\Ftrans}{\mathcal{F}}
% Experiments
\newcommand{\omavg}{\omega_0}
\newcommand{\omsig}{\sigma_{\omega}}
\newcommand{\Dphi}{\Delta \phi}

@@

We now introduce the truncated KdV (TKdV) system, which is the main focus of this study. Consider the Fourier series of the state variable
\begin{align}
&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k(t) \, e^{i k x} \, , \\
\label{uhat}
&\uhat_k(t) = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x,t) \, e^{-i k x} \dx
\end{align}
Since $u$ is real valued, $\uhat_{-k} = \uhat_{k}^*$, and since momentum vanishes $\uhat_0 = 0$.
Next, consider the Galerkin truncation at wave number $\Lambda$
\begin{align}
\uL(x,t) = \Proj u =
\sum_{\abs{k} \le \Lambda} \uhat_k(t) \, e^{i k x} \, , \qquad
\end{align}
where $\Proj$ is a projection operator and \eqref{uhat} still holds. Projecting the KdV equation \eqref{dimlessKdV} onto the finite dimensional space gives the truncated KdV equation (TKdV)
\begin{align}
\label{TKdV}
&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad
C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2}
\end{align}
Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$. The constants $C_3$ and $C_2$ are the same as before and have been repeated here for convenience.

@@

% NOTES
%Further, since the regime is shallow, short-time dispersive effects are minimal, meaning that all waves travel at nearly the same speed $c = \sqrt{g \depth}$, independent of wavelength.

@@

	The numerical discretization first separates the Hamiltonian into linear and nonlinear components. Then a pseudo-spectral scheme is used for spatial discretization, with de-aliasing applied to the nonlinear term $u^2$

@@

\begin{align}
\label{TKdV2}
&\pd{\uL}{t} = F = 
-  \frac{1}{2} C_3 \drat^{-3/2} \,\Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0 \qquad \text{for } x \in [-\pi,\pi]
\end{align}

@@

A pseudo-spectral scheme is used for spatial discretization. Proper de-aliasing methods should be used to calculate the nonlinear term $u^{2}$ in the equation. Here the standard 2/3-rule is applied to calculate the quadratic term in physical domain and extend the spectral modes by adding additional zeros in the tails. That is, we explicitly calculate the spectral mode of the square term by 
\[
\widehat{\left(u^{2}\right)}_{k} = \Ftrans \left[U^{2}\right],\quad U 
= \Ftrans^{-1} \left[\left\{ \hat{u}_{k}\mid_{\left|k\right|\leq\frac{\Lambda}{2}},0\mid_{\frac{\Lambda}{2}<\left|k\right|\leq\frac{3}{2}\Lambda}\right\} \right],
\]
where $\Ftrans$ represents the operator for Fourier transform.
Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\} _{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ F_{k}\right\}$
\[
\dot{\hat{u}}_{k}=F_{k}\left(u\right)=C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
\]
Different scales are mixed due to the nonlinear term $u^{2}$. The linear term including third derivative might become stiff when the total wavenumbers become large.

The symplectic scheme
used here for the time integration of the TKdV equation is the 4th-order midpoint method \cite{mclachlan1993symplectic}.
Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$ the
discretized operator for the spectral modes. The midpoint method has
two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$ so that
\[
\begin{aligned}\mathbf{y}_{1}-\mathbf{u}^{n} & =w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
\mathbf{y}_{2}-\mathbf{y}_{1} & =w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
\mathbf{u}^{n+1}-\mathbf{y}_{2} & =w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
\end{aligned}
\]
with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$. Then the PDFs can be counted
from the solution $u_{\Lambda}\left(t\right)$, and the statistical
expectations can be calculated by averaging along the trajectory
\[
\left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
\]
given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
already reaching the stationary state.

For the symplectic schemes, implicit iteration method is required to get the accurate solution. For any implicit problem of the form
\[
\mathbf{z}-\mathbf{u}^{n}=\tau\mathbf{F\left(z\right)}=\tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
\]
we split the operators into a linear and nonlinear components and only treat the nonlinear part on the right hand side for the previous stage in the iteration. Then we have the iterative scheme for the solution of implicit problem 
\[
\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]}=\mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
\]
The linear operator $\mathbf{L}$ is the diagonal differential operator, thus it is easy to invert in these schemes. The termination criteria is determined by
\[
\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert <\delta\;\mathrm{and}\;\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert \geq\left\Vert \mathbf{z}^{\left[l\right]}-\mathbf{z}^{\left[l-1\right]}\right\Vert .
\]
It is found that the iteration tolerance can be set as $\delta=1\times10^{-10}$.
Finally, the starting guess for the iterations can be found by quadratic extrapolation for the previous three stages $\left(t_{u},\mathbf{u}^{n}\right), \left(t_{u}-\tau,\mathbf{u}^{n-1}\right), \left(t_{u}-2\tau,\mathbf{u}^{n-2}\right)$. Therefore we can find the extrapolated value at $t=t_{n}+\beta\tau$ by
\[ \mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
\]

>>>>

%----------------------------------------------------------------------------%
%% LATEX DEFINITIONS
%----------------------------------------------------------------------------%
% Basic editing
\newcommand{\tocite}{{\color{blue}(to cite)}}
\newcommand{\vsp}[1]{\vspace{#1 pc} \noindent}
\newcommand{\np}{\newpage \noindent}
\newcommand{\nick}[1]{{\color{red} #1}}
% Derivatives
\newcommand{\pd}[2]    { \frac{\partial #1} {\partial #2} }
\newcommand{\ppd}[2]  { \frac{\partial^2 #1}{{\partial #2}^2} }
\newcommand{\pdi}[2] { {\partial_#2} #1 }
\newcommand{\td}[2] { \frac{d #1} { d #2 } }
\newcommand{\grad}{\nabla}
\newcommand {\Lap} {\grad^2}
% Vectors and operators
\newcommand{\bvec}[1]{\ensuremath{\boldsymbol{#1}}}
\newcommand{\abs}[1]{\left| #1 \right|}
\newcommand{\norm}[1]{\left\| #1 \right\|}
\newcommand{\mean}[1]{\left< #1 \right>}
\newcommand{\eps}{\varepsilon}
\newcommand{\dx}{\, dx}
\newcommand{\CC}{{\mathbb{C}}}

% Basic physical parameters and scales
\newcommand{\freqp}{f_p}
\newcommand{\etastd}{\eta_{\text std}}
% Parameters that change crossing the ADC
\newcommand{\depth}{d}
\newcommand{\dup}{\depth_{-}}
\newcommand{\ddn}{\depth_{+}}
\newcommand{\dupdn}{\depth_{\pm}}
\newcommand{\lam}{\lambda}
\newcommand{\lamup}{\lam_{-}}
\newcommand{\lamdn}{\lam_{+}}
\newcommand{\lamupdn}{\lam_{\pm}}
\newcommand{\lamfac}{N}
\newcommand{\drat}{\mathcal{D}}
\newcommand{\dratdn}{\drat_+}
\newcommand{\dratupdn}{\drat_{\pm}}
% Other experimental stuff
\newcommand{\omavg}{\omega_0}
\newcommand{\omsig}{\sigma_{\omega}}
\newcommand{\Dphi}{\Delta \phi}
% Statistical quantities
\newcommand{\En}{\mathcal{E}}
\newcommand{\Mo}{\mathcal{M}}
\newcommand{\skw}{\text{skew}}
\newcommand{\skwdn}{\skw_+}
\newcommand{\var}{\text{var}}
\newcommand{\varup}{\var_-}
\newcommand{\vardn}{\var_+}
\newcommand{\varupdn}{\var_{\pm}}
\newcommand{\kurt}{\text{kurt}}
\newcommand{\std}{\text{std}}

% Dimensionless parameters
\newcommand{\ampscale}{\mathcal{A}}
\newcommand{\lengthscale}{\mathcal{L}}
\newcommand{\timescale}{\mathcal{T}}
\newcommand{\epsup}{\eps_0}
\newcommand{\delup}{\delta_0}
% Hamiltonian stuff
\newcommand{\uhat}{\hat{u}}
\newcommand{\sympJ}{\mathcal{J}}
\newcommand{\vard}[2]{\frac{\delta #1}{\delta #2}}
\newcommand{\Ham}{\mathcal{H}}
\newcommand{\Hup}{\Ham^{-}}
\newcommand{\Hdn}{\Ham^{+}}
\newcommand{\Hupdn}{\Ham^{\pm}}
\newcommand{\Hthree}{\Ham_{3}}
\newcommand{\Htwo}{\Ham_{2}}
\newcommand{\Fcnl}{\mathcal{F}}
% Truncated stuff
\newcommand{\Proj}{\mathcal{P}_{\Lambda}}
\newcommand{\uL}{u_{\Lambda}}
\newcommand{\HLupdn}{\Ham_{\Lambda}^{\pm}}
\newcommand{\SympL}{\sympJ_{\Lambda}}
\newcommand{\uup}{u_{-}}
\newcommand{\udn}{u_{+}}
\newcommand{\uupdn}{u_{\pm}}
% Gibbs and theta
\newcommand{\Gibbs}{\mathcal{G}}
\newcommand{\Gup}{\Gibbs^{-}}
\newcommand{\Gdn}{\Gibbs^{+}}
\newcommand{\Gupdn}{\Gibbs^{\pm}}
\newcommand{\thup}{\theta^{-}}
\newcommand{\thdn}{\theta^{+}}
\newcommand{\thupdn}{\theta^{\pm}}
\newcommand{\meanup}[1]{\mean{#1}_{-}}
\newcommand{\meandn}[1]{\mean{#1}_{+}}
\newcommand{\meanupdn}[1]{\mean{#1}_{\pm}}
\newcommand{\Gz}{\Gibbs_0}
\newcommand{\meanz}[1]{\mean{#1}_0}
% Misc stuff
\newcommand{\tavg}[1]{\overline{#1}}
\newcommand{\transf}{\mathcal{F}}
\newcommand{\Nsamp}{N_s}
\newcommand{\sumsamp}{\sum_{i=1}^{\Nsamp}}
\newcommand{\traveldist}{\ell}
\newcommand{\Ftrans}{\mathcal{F}}
\newcommand{\RHS}{F}
\newcommand{\RHSh}{\hat{\RHS}}
\newcommand{\usqdah}{ (\widehat{u_{da}^2}) }

@@

We now introduce the truncated KdV (TKdV) system, which is the main focus of this study. Consider the Fourier series of the state variable
\begin{align}
&u(x,t) = \sum_{k=-\infty}^{\infty} \uhat_k(t) \, e^{i k x} \, , \\
\label{uhat}
&\uhat_k(t) = \frac{1}{2 \pi} \int_{-\pi}^{\pi} u(x,t) \, e^{-i k x} \dx
\end{align}
Since $u$ is real valued, $\uhat_{-k} = \uhat_{k}^*$, and since momentum vanishes $\uhat_0 = 0$.
Next, consider the Galerkin truncation at wave number $\Lambda$
\begin{align}
\uL(x,t) = \Proj u =
\sum_{\abs{k} \le \Lambda} \uhat_k(t) \, e^{i k x} \, , \qquad
\end{align}
where $\Proj$ is a projection operator and \eqref{uhat} still holds. Projecting the KdV equation \eqref{dimlessKdV} onto the finite dimensional space gives the truncated KdV equation (TKdV)
\begin{align}
\label{TKdV}
&\pd{\uL}{t} +  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 + C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = 0
\qquad \text{for } x \in [-\pi,\pi] \\
&C_3 = \frac{3}{2} \pi^{1/2} \epsup \delup^{-1} \, , \quad C_2 = \frac{2 \pi^2 \delup}{3 \lamfac^2}
\end{align}
Note the additional projection operator in front of the quadratic term $\uL^2$, which removes the aliased modes of wavenumber larger than $\Lambda$. The constants $C_3$ and $C_2$ are the same as before and have been repeated here for convenience.

@@

@@

Rearranging the TKdV system \eqref{TKdV} gives
\begin{align}
\label{TKdV2}
&\pd{\uL}{t} = -  \frac{1}{2} C_3 \drat^{-3/2} \, \pd{}{x} \Proj (\uL)^2 - C_2 \drat^{1/2} \, \frac{\partial^3 \uL}{\partial x^3} = \RHS
%\qquad \text{for } x \in [-\pi,\pi]
\end{align}	
where we have represented the right-hand side by $\RHS$.
Our numerical scheme first separates the Hamiltonian into linear and nonlinear components. Then a pseudo-spectral scheme is used for spatial discretization with de-aliasing applied to the nonlinear term $\Proj (\uL)^2$ according to the standard 2/3-rule. That is, we first pad $\uhat_k$ with zeros for $\Lambda < \abs{k} \le 3 \Lambda$ \nick{3 or 3/2???}, transform to physical space and square to obtain $\uL^2$ on a fine grid, then transform back to spectral space and truncate to obtain the Fourier coefficients of $\Proj (\uL)^2$
For simplicity, we denote these Fourier coefficients $\usqdah_k$. That is,
\begin{equation}
\Proj (\uL)^2 = \sum_{\abs{k} \le \Lambda} \usqdah_k e^{i k x}
\end{equation}

@@

@@

Therefore the ODE system after spatial discretization can be written for the spectral vector $\mathbf{u}=\left\{ \hat{u}_{k}\right\}_{k=0}^{\Lambda}$ and $\mathbf{F}=\left\{ \RHSh_{k} \right\}_{k=0}^{\Lambda}$
\begin{equation}
\dot{\hat{u}}_{k} = \RHSh_k \left[u\right] = C_{2}\mathcal{D}^{1/2}ik^{3}\hat{u}_{k}-\frac{1}{2}C_{3}\mathcal{D}^{-3/2}ik\widehat{\left(u^{2}\right)}_{k}.
\end{equation}
Different scales are mixed due to the nonlinear term $u^{2}$. The linear term including third derivative might become stiff when the total wavenumbers become large.

The symplectic scheme
used here for the time integration of the TKdV equation is the 4th-order midpoint method \cite{mclachlan1993symplectic}.
Let $\mathbf{u}^{n}$ be the solution at time $t_{n}$ and $\mathbf{u}^{n+1}$
be the solution at next time step $t_{n+1}=t_{n}+\Delta t$, and $\mathbf{F}$ the
discretized operator for the spectral modes. The midpoint method has
two intermediate stages and the auxiliary vectors $\mathbf{y}_{1},\mathbf{y}_{2}$ so that
\begin{align}
&\mathbf{y}_{1}-\mathbf{u}^{n} = w_{1}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{1}+\mathbf{u}^{n}\right]\right),\\
&\mathbf{y}_{2}-\mathbf{y}_{1} = w_{2}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{y}_{2}+\mathbf{y}_{1}\right]\right),\\
&\mathbf{u}^{n+1}-\mathbf{y}_{2} = w_{3}\Delta t\mathbf{F}\left(\frac{1}{2}\left[\mathbf{u}^{n+1}+\mathbf{y}_{2}\right]\right),
\end{align}
with the time increments $w_{1}=\left(2+2^{1/3}+2^{-1/3}\right)/3$,
$w_{2}=1-2w_{1}$, and $w_{3}=w_{1}$. Then the PDFs can be counted
from the solution $u_{\Lambda}\left(t\right)$, and the statistical
expectations can be calculated by averaging along the trajectory
\begin{equation}
\left\langle F\left(u_{\Lambda}\right)\right\rangle _{\mu}=\frac{1}{T}\int_{t_{0}}^{t_{0}+T}F\left(u_{\Lambda}\left(t\right)\right)dt,
\end{equation}
given ergodicity in the solution with $u_{\Lambda}\left(t_{0}\right)$
already reaching the stationary state.

For the symplectic schemes, implicit iteration method is required to get the accurate solution. For any implicit problem of the form
\begin{equation}
\mathbf{z}-\mathbf{u}^{n}=\tau\mathbf{F\left(z\right)}=\tau\mathbf{Lz}+\tau\mathbf{N\left(z\right)},
\end{equation}
we split the operators into a linear and nonlinear components and only treat the nonlinear part on the right hand side for the previous stage in the iteration. Then we have the iterative scheme for the solution of implicit problem 
\begin{equation}
\left(1-\tau\mathbf{L}\right)\mathbf{z}^{\left[l+1\right]}=\mathbf{u}^{n}+\tau\mathbf{N}\left(\mathbf{z}^{\left[l\right]}\right).
\end{equation}
The linear operator $\mathbf{L}$ is the diagonal differential operator, thus it is easy to invert in these schemes. The termination criteria is determined by
\begin{equation}
\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert <\delta\;\mathrm{and}\;\left\Vert \mathbf{z}^{\left[l+1\right]}-\mathbf{z}^{\left[l\right]}\right\Vert \geq\left\Vert \mathbf{z}^{\left[l\right]}-\mathbf{z}^{\left[l-1\right]}\right\Vert .
\end{equation}
It is found that the iteration tolerance can be set as $\delta=1\times10^{-10}$.
Finally, the starting guess for the iterations can be found by quadratic extrapolation for the previous three stages $\left(t_{u},\mathbf{u}^{n}\right), \left(t_{u}-\tau,\mathbf{u}^{n-1}\right), \left(t_{u}-2\tau,\mathbf{u}^{n-2}\right)$. Therefore we can find the extrapolated value at $t=t_{n}+\beta\tau$ by
\begin{equation}
\mathbf{u}^{n+\beta}=\mathbf{Q}\left(t_{n}+\beta\tau\right)=\left[\frac{1}{2}\beta\left(\beta+3\right)+1\right]\mathbf{u}^{n}-\beta\left(\beta+2\right)\mathbf{u}^{n-1}+\frac{1}{2}\beta\left(\beta+1\right)\mathbf{u}^{n-2}.
\end{equation}
